Secure Composition \\ of Robust and Optimising Compilers \vspace{-1.5em}

### 2024-05-29 csf2/main.tex
nit

  \item %
  		This paper provides formal proofs of the security properties preserved by each of the presented passes. 
  		% 
  		Additionally this paper analyzes each pass in terms of their compatibility with each other (\Cref{sec:formalities}), proving how does each pass fulfill the conditions that make their composition meaningful from a security standpoint.
        % In doing so, the paper provides formal insights into some technicalities of the concrete proofs necessary as technical setup to reap the benefits of the presented compositionality framework.

@@

To analyze the security of compilation passes and their interaction within a compilation pipeline, we rely on a few key notions: the definition of a trace relation, and the definition of when is a trace relation well-formed with respect to a class.
%
Consider traces $\src{\trace}$ and $\trg{\trace}$ as well as two trace relations $\sim_1$ and $\sim_2$. 
% 
The traces are related $\src{\trace}\sim_1\bullet\sim_2\trg{\trace}$ if there is another trace $\irl{\trace}$ such that $\src{\trace}\sim_1\irl{\trace}$ and $\irl{\trace}\sim_2\trg{\trace}$.
% 
\bul{A relation $\sim$ is well-formed w.r.t. a class of target-level properties $\trg{\class}$} iff \rul{the universal image preserves set membership}.
\begin{definition}[Well-formedness of $\sim$ for a Class $\trg{\class}$]\label{def:wfc:sig:tracerel}
  % $$

@@

The trace model is extended with $\ird{{Spec}}$, $\ird{{Rlb}}$, and $\ird{{Barrier}}$ events that signal the respective operational action. 
A $\ird{\lbarrier}$ prevents any execution whatsoever besides $\ird{{Rlb}}$ when run in speculative execution mode and does nothing when run in normal mode, e.g., \Cref{tr:e-spec-barrier}.%
\footnote{In the rule, the notation $\cfstate' = \cfstate[.\text{win} = 0]$ means that $\cfstate'$ is a copy of $\cfstate$ up to field $\text{win}$, which is set to $0$.}
Lastly, $\ird{Rlb}$ is emitted when the speculation window is zero and the operationl state is rolled back (see \Cref{tr:e-spec-eaten}).

@@

% \Cref{thm:rtpsim:sig} is applicable, up to the well-formedness of $\sim_1$ w.r.t. $\trg{\class[_2]}$.
Consider applying $\ccspec$ first and then $\ccb$ (albeit currently syntactically impossible).
% , let us pretend that it is possible by just extending $\ccb$ for the constructs present in $\Lscct$ that are not present in $\Ltrg$.
% Clearly, as evidenced earlier (\Cref{subsec:formalities:props}), $\ccb\cdot\ccspec$ is an acceptable composition that yields an interesting class of security properties, i.e., $\sigma_{\sim_{\irl{ms}}\bullet\sim_{\ird{\mathghost}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$. 
In this case, $\ccb$ would insert new branches into the code that are not protected by a speculation barrier! 
This concern is reflected in the proofs that establish that the security class resulting of the composition of the trace relations is meaningful.
% , much similar to ensuring that a definition models what the formal methods expert intends to model. 
For the $\ccspec\cdot\ccb$ case, the class is $\sigma_{\sim_{\ird{\mathghost}}\bullet\sim_{\irl{ms}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$.
Since $\ev{Use\ \loc\ n}\sim_{\irl{ms}}\irl{Branch\ n}\cdot\irl{Spec}\cdots$, where $\cdots$ does not contain a $\ev{{Barrier}}$ event, the resulting class is \emph{not} the original $\sssafe$ that is intended and it would break the corresponding \Cref{thm:prop-rel-corr}.
% Because of this, it is crucial to analyze the precise shape of the class of properties after mapping a target-level property to a source-level property\footnote{Dito for mapping a source-level property to a target-level property.}.

@@

The robust preservation of properties as a compiler-level criterion has been analyzed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2019survey} and thus we build on that framework.
No existing work is concerned with composing robustly safe compilers.
%Parts of these works consider languages with different trace models and our technical setup can handle this. 
The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.
The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:sequential}) but use it in the context of protocols.
The work does not consider the generality to support different trace models.
% The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:sequential}) as well as in an earlier version of this work~\cite{kruse2022csc}.
% However, they do not demonstrate the scalability of the approach by means of a case study.

@@

To extend the languages in this paper with a less restricted form of pointer arithmetic, the region coloring memory safety monitor presented in earlier work~\cite{michael2023mswasm} can be used.
The work presenting this monitor provides an approach for the robust preservation of memory safety compiling from C to WASM.
However, they do not discuss composition of secure compilers but rather investigate an instance of a secure compiler.

>>>>

  \item %
  		This paper provides formal proofs of the security properties preserved by each of the presented passes. 
  		% 
  		Additionally this paper analyses each pass in terms of their compatibility with each other (\Cref{sec:formalities}), proving how does each pass fulfill the conditions that make their composition meaningful from a security standpoint.
        % In doing so, the paper provides formal insights into some technicalities of the concrete proofs necessary as technical setup to reap the benefits of the presented compositionality framework.

@@

To analyse the security of compilation passes and their interaction within a compilation pipeline, we rely on a few key notions: the definition of a trace relation, and the definition of when is a trace relation well-formed with respect to a class.
%
Consider traces $\src{\trace}$ and $\trg{\trace}$ as well as two trace relations $\sim_1$ and $\sim_2$. 
% 
The traces are related $\src{\trace}\sim_1\bullet\sim_2\trg{\trace}$ if there is another trace $\irl{\trace}$ such that $\src{\trace}\sim_1\irl{\trace}$ and $\irl{\trace}\sim_2\trg{\trace}$.
% 
\bul{A relation $\sim$ is well-formed w.r.t. a class of target-level properties $\trg{\class}$} iff \rul{the universal image preserves set membership}.
\begin{definition}[Well-formedness of $\sim$ for a Class $\trg{\class}$]\label{def:wfc:sig:tracerel}
  % $$

@@

The trace model is extended with $\ird{{Spec}}$, $\ird{{Rlb}}$, and $\ird{{Barrier}}$ events that signal the respective operational action. 
A $\ird{\lbarrier}$ prevents any execution whatsoever besides $\ird{{Rlb}}$ when run in speculative execution mode and does nothing when run in normal mode, e.g., \Cref{tr:e-spec-barrier}.%
\footnote{In the rule, the notation $\cfstate' = \cfstate[.\text{win} = 0]$ means that $\cfstate'$ is a copy of $\cfstate$ up to field $\text{win}$, which is set to $0$.}
Lastly, $\ird{Rlb}$ is emitted when the speculation window is zero and the operational state is rolled back (see \Cref{tr:e-spec-eaten}).

@@

% \Cref{thm:rtpsim:sig} is applicable, up to the well-formedness of $\sim_1$ w.r.t. $\trg{\class[_2]}$.
Consider applying $\ccspec$ first and then $\ccb$ (albeit currently syntactically impossible).
% , let us pretend that it is possible by just extending $\ccb$ for the constructs present in $\Lscct$ that are not present in $\Ltrg$.
% Clearly, as evidenced earlier (\Cref{subsec:formalities:props}), $\ccb\cdot\ccspec$ is an acceptable composition that yields an interesting class of security properties, i.e., $\sigma_{\sim_{\irl{ms}}\bullet\sim_{\ird{\mathghost}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$. 
In this case, $\ccb$ would insert new branches into the code that are not protected by a speculation barrier! 
This concern is reflected in the proofs that establish that the security class resulting of the composition of the trace relations is meaningful.
% , much similar to ensuring that a definition models what the formal methods expert intends to model. 
For the $\ccspec\cdot\ccb$ case, the class is $\sigma_{\sim_{\ird{\mathghost}}\bullet\sim_{\irl{ms}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$.
Since $\ev{Use\ \loc\ n}\sim_{\irl{ms}}\irl{Branch\ n}\cdot\irl{Spec}\cdots$, where $\cdots$ does not contain a $\ev{{Barrier}}$ event, the resulting class is \emph{not} the original $\sssafe$ that is intended and it would break the corresponding \Cref{thm:prop-rel-corr}.
% Because of this, it is crucial to analyse the precise shape of the class of properties after mapping a target-level property to a source-level property\footnote{Dito for mapping a source-level property to a target-level property.}.

@@

The robust preservation of properties as a compiler-level criterion has been analysed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2019survey} and thus we build on that framework.
No existing work is concerned with composing robustly safe compilers.
%Parts of these works consider languages with different trace models and our technical setup can handle this. 
The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.
The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:sequential}) but use it in the context of protocols.
The work does not consider the generality to support different trace models.
% The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:sequential}) as well as in an earlier version of this work~\cite{kruse2022csc}.
% However, they do not demonstrate the scalability of the approach by means of a case study.

@@

To extend the languages in this paper with a less restricted form of pointer arithmetic, the region colouring memory safety monitor presented in earlier work~\cite{michael2023mswasm} can be used.
The work presenting this monitor provides an approach for the robust preservation of memory safety compiling from C to WASM.
However, they do not discuss composition of secure compilers but rather investigate an instance of a secure compiler.
